---
title: "A Guide to rsync"
author: "Nick Hoffman"
date: "September 2024"
output:
  html_document:
    df_print: paged
    highlight: pygment
    keep_md: no
    toc: true
    number_sections: true
    toc_depth: 1
    toc_float: true
    theme: journal
editor_options:
    chunk_output_type: inline
---

<style type="text/css">
h2, h3, h4, h5, h6 {
  counter-reset: section;
}
p {
  font-size:18px;
}

ul {
  font-size:18px;
}

li {
  font-size:18px;
}
table {
   padding: 0;border-collapse: collapse;
   layout: fixed;
   width: 90%; }
table tr {
   border-top: 1px solid #cccccc;
   background-color: white;
   margin: 0;
   padding: 0; }
table tr:nth-child(2n) {
   background-color: #f8f8f8; }
table tr th {
   font-weight: bold;
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr td {
   border: 1px solid #cccccc;
   margin: 0;
   padding: 6px 13px; }
table tr th :first-child, table tr td :first-child {
   margin-top: 0; }
table tr th :last-child, table tr td :last-child {
   margin-bottom: 0; }
.html-widget {
    margin: auto;
}
</style>

---

```{r, results='asis', echo = F}
toc_depth <- rmarkdown::metadata$output$html_document$toc_depth
sel <- paste0("h",(toc_depth+1):10, collapse = " > span, ")
cat(paste0("<style>",
           sel, 
           " > .header-section-number { display: none; } </style>"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment=NA)
```


# Introduction

This guide is intended to provide some minimum information on the software tool rsync for INSITE's data producers, so they can send large files remotely to our file server. This guide is a quick start. If you would like more information about advanced rsync use, please consult this <a href="https://vamsipenmetsa.medium.com/efficient-syncing-with-rsync-command-2b5a9dd9e13c" target="_blank">page</a>.

# What is rsync?

rsync (<i>remote sync</i>) is a software tool for transferring and synchronizing large files across different computer systems. It works incrementally, by comparing and correcting only the differences between two versions of a file, rather than recopying an entire file. Computers with Linux operating systems (e.g., our high performance computing clusters), as well as macOS, tend to support rsync by default. Computers with Windows operating systems can be configured in order to support rsync.

# How do I use rsync?

rsync commands must be input into a command-line interface (CLI) like Terminal or Command Prompt. rsync involves syncing files between two different computer systems, which must both be rsync-compatible in order for the operation to succeed. The INSITE server is rsync-compatible, and if the computer from which you're transferring files uses Linux or Mac operating system, it should be rsync-compatible too. If you are using a Windows operating system, see below for more information.  

Although the INSITE server is rsync-compatible, by default the rsync functionality is turned off for users. In order to use rsync with the INSITE server, you must reach out to the INSITE data manager, Nick Hoffman, and request rsync permissions. He will provide you with an rsync password which you will need in the course of sending data through rsync to the server.  

To send a file from a local computer to a remote one, the following syntax is appropriate. (Note that anything in brackets needs to be customized depending on your particular use case.)

```{r rsync-syntax, include=TRUE,echo=FALSE}
cat(paste("rsync [optionalFlags] [pathToLocalFile] [pathToRemoteFolder]"))
```

In our case, the remote folder to which we will be sending our files is the INSITE server, so we can replace 

```{r rsync-example, include=TRUE,echo=FALSE} 
cat(paste("[pathToRemoteFolder]"))
```

with 

```{r rsync-example2, include=TRUE,echo=FALSE} 
cat(paste("[Username]@[ServerIPAddress]::[folder]."))
```

The server IP address is available upon request from Nick Hoffman, and our shared folder for INSITE data is  


```{r rsync-example6, include=TRUE,echo=FALSE} 
cat(paste("insiteData/data/...."))
```


Your username is specific to you. The rsync syntax may then look like:

```{r rsync-syntax2,include=TRUE,echo=FALSE}
cat(paste("rsync [optionalFlags] [pathToLocalFile] NickHoffman@[ServerIPAddress]::insiteData/data/leafhopper/theme2/genome/exp5/"))

```

If you're performing your rsync from within the folder in which reside the data you intend to send, then you simply need to replace

```{r rsync-exampl3, include=TRUE,echo=FALSE}

cat(paste('[pathToLocalFile]' ))
```

with the name of your file. If you are performing rsync from a distinct folder, you need to input the path from the folder you're currently in to the location of your file. For the sake of simplicity, let's assume we're sending from the folder in which our data reside, and the name of our genomic leafhopper data file is 

```{r rsync-exampl4, include=TRUE,echo=FALSE}

cat(paste('Hoffman_09272024_3.fastq.' ))
```

In that case, you may perform the following command, omitting for now the [optionalFlags]: 

```{r rsync-syntax3, include=TRUE,echo=FALSE}
cat(paste("rsync Hoffman_09272024_3.fastq NickHoffman@[ServerIPAddress]::insiteData/data/leafhopper/theme2/genome/exp5/"))
```

Your terminal should then request your password. Note that the password it requires is not your regular sign-in password for the server, but the particular rsync password mentioned above that will be provided to you by Nick Hoffman.

## rsync flags

But what about those optional flags? There are <a href="https://ss64.com/bash/rsync_options.html" target="_blank">many flags</a> which you could choose to incorporate into your rsync command.

Here are some flags we recommend using in your rsync command.

1. In order for rsync to proceed as quickly as possible, we recommend using the flag *-z*, which compresses file data as they are sent.

2. If you intend to transfer a whole folder, rather than just a file, we recommend using the flag *-a*, which specifies that rsync should proceed recursively (that is, rsync the folder, subfolders, their files, etc.).

3. If rsync is not succeeding, you can add the flag *-v* for a verbose output. This will cause your machine to log more information about what's going on under the hood, which may enable you to troubleshoot your error.

This could look as follows:

```{r rsync-syntax6, include=TRUE,echo=FALSE}
cat(paste("rsync -avz Hoffman_09272024_3.fastq NickHoffman@[ServerIPAddress]::insiteData/data/leafhopper/theme2/genome/exp5/"))
```

You are of course encouraged to explore other flags linked above in the advanced user guide.

# What if I'm on Windows?

If you're on Windows, rsync won't be native to your operating system. However, you can download rsync through <a href="https://www.cygwin.com/install.html" target="_blank">the Cygwin Project</a>. Please follow the guide there in order to download rsync, and reach out to Nick Hoffman if you have any problem with it. If you're not sure whether rsync downloaded properly onto your Windows system, you can run the following command in <b>Cygwin Terminal</b>:

```{r rsync-syntax4, include=TRUE,echo=FALSE}
cat(paste("rsync --version"))
```

If rsync installed correctly, you should should see an output like:


```{r rsync-5, include=TRUE,echo=FALSE}
cat(paste("rsync  version x.x.x  protocol version xxx"))
```


To navigate in the Cygwin Terminal to where your files are located, you'll need to add

``` {r cyg, include=TRUE,echo=FALSE}

cat(paste("/cygdrive/c/"))
```

in front of your regular directory. So to navigate to my documents folder in Cygwin Terminal, I write

``` {r cyg-2, include=TRUE, echo=FALSE}

cat(paste("cd /cygdrive/c/Users/Nick/Documents"))
```


and I can rsync my files from there.  
    
Lastly, because Synology software enforces a stricter set of permissions with respect to Windows/Cygwin than with other operating systems, you'll need to write directly to your target file, rather than renaming a temporary file. To achieve this, use the *\-\-inplace* flag. For example:


```{r rsync-syntax100, include=TRUE,echo=FALSE}
cat(paste("rsync -avz --inplace Hoffman_09272024_3.fastq NickHoffman@[ServerIPAddress]::insiteData/data/leafhopper/theme2/genome/exp5/"))
```

Happy syncing !
